---
title: "Challenge 1 solution"
author: "Sidharth"
desription: "Something to describe what I did"
date: "06/16/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge1
  - Sid
  - dataset
  - ggplot2
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

This document provides yaml header inforamtion you will need to replicate each week to submit your homework or other blog posts. Please observe the following conventions:

-   Save your own copy of this template as a blog post in the `posts` folder, naming it `FirstLast_hwX.qmd`
-   Edit the yaml header to change your author name - use the same name each week
-   include a *description* that is reader friendly
-   update the *category* list to indicate the type of submission, the data used, the main packages or techniques, your name, or any thing else to make your document easy to find
-   edit as a normal qmd/rmd file


```{r}
x <- c(2,3,4,5)
mean(x)
```

## Rendering your post

I did this