---
title: "UFO Shapes & Length of Encounters: Descriptive Analysis"
author: "Peri (Teresa) Lardo"
desription: "Descriptive analysis of UFO sightings data set"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Project1
  - Peri Teresa Lardo
  - UFO sightings
---

## Load libraries and dataset, clean data

```{r}
#| label: setup
#| warning: false

#Load libraries and import data
library(tidyverse)
library(readr)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)

# Create shape categories for personal sanity

round <- c("disk", "round", "sphere", "circle", "dome")
triangular <- c("triangle", "cone", "pyramid", "delta", "chevron")
oval <- c("oval", "teardrop", "egg", "crescent")
parallelogram <- c("cross", "diamond", "hexagon", "rectangle")
cylinder <- c("cylinder", "cigar")


ufo_sighting_data <- read_csv("data/ufo_sighting_data.csv") %>% drop_na(UFO_shape, length_of_encounter_seconds) %>%
  filter(!UFO_shape %in% c("light", "fireball", "flare", "flash", "other", "unknown", "formation", "changing", "changed")) %>%
  mutate(Shape_Category = case_when(
         UFO_shape %in% round ~ "Round",
         UFO_shape %in% triangular ~ "Triangular",
         UFO_shape %in% oval ~ "Oval",
         UFO_shape %in% parallelogram ~ "Parallelogram",
         UFO_shape %in% cylinder ~ "Cylinder"))
```
Since I'm looking into UFO shapes, and there are a lot of individual shapes, I opted to first remove "shapes" that aren't literally shapes (including descriptions like "flash" or "formation" or "changing" as well as the super-specific shape of "light"). Then I collapsed the remaining UFO shape descriptions into 5 categories: Triangular, Cylindrical, Oval, Round (more properly circular than oval), and Parallelograms.

## Probability Distributions of Encounter Length

```{r}
# Plot the probability & CDF for variables
probability <- ufo_sighting_data$length_of_encounter_seconds
plot(probability,
     xlab = "Length of UFO Encounter in Seconds",
     main = "Probability",
     col = "navyblue", 
     cex = 0.8)
```

```{r}
cum_probability <- cumsum(probability)
plot(cum_probability,
     xlab = "Length of UFO Encounter in Seconds",
     ylab = "Cumulative Probability",
     main = "Cumulative Probability Distribution",
     col = "red",
     cex = 0.8)
```

## Grouping Data by Shape Category

```{r}
# Group data by the 2 variables of interest. Get mean & STDEV for at least 1 variable and print results
round <- c("disk", "round", "sphere", "circle", "dome")
triangular <- c("triangle", "cone", "pyramid", "delta", "chevron")
oval <- c("oval", "teardrop", "egg", "crescent")
parallelogram <- c("cross", "diamond", "hexagon", "rectangle")
cylinder <- c("cylinder", "cigar")

avgs <- ufo_sighting_data %>%
  group_by(Shape_Category) %>%
  summarize(mean(length_of_encounter_seconds),
            sd(length_of_encounter_seconds),
            n())

colnames(avgs) <- c("Shape", "mean_seconds", "sd_seconds", "n_seconds")

print(avgs)
```
## Split Variables

```{r}
# Split your variables
cylindrical_UFOS <- avgs %>% dplyr::filter(Shape == "Cylinder")
oval_UFOS <- avgs %>% dplyr::filter(Shape == "Oval")
parallelogram_UFOS <- avgs %>% dplyr::filter(Shape == "Parallelogram")
round_UFOS <- avgs %>% dplyr::filter(Shape == "Round")
triangular_UFOs <- avgs %>% dplyr::filter(Shape == "Triangular")
```

```{r}
# Rename columns according to the splits you made; Obtain difference in means, standard errors, and confidence intervals
colnames(cylindrical_UFOS) <- c("Shape", "Y_bar_cy", "sd_cy", "n_cy") 

colnames(oval_UFOS) <- c("Shape", "Y_bar_ov", "sd_ov", "n_ov")

colnames(parallelogram_UFOS) <- c("Shape", "Y_bar_pa", "sd_pa", "n_pa")

colnames(round_UFOS) <- c("Shape", "Y_bar_ro", "sd_ro", "n_ro")

colnames(triangular_UFOs) <- c("Shape", "Y_bar_tr", "sd_tr", "n_tr")

```

## Estimate Mean Differences, Standard Errors & CIs for Triangular v. Round UFOs

```{r}
# Gaps for Triangular versus Round UFOs

gap_tro <- triangular_UFOs$Y_bar_tr - round_UFOS$Y_bar_ro

gap_tro_se <- sqrt((triangular_UFOs$sd_tr ^ 2 / triangular_UFOs$n_tr) + (round_UFOS$sd_ro ^ 2 / round_UFOS$n_ro))

gap_tro_ci_low <- gap_tro - 1.96 * gap_tro_se

gap_tro_ci_up <- gap_tro + 1.96 * gap_tro_se

result_tro <- cbind(triangular_UFOs[,-1], round_UFOS[,-(1:2)], gap_tro, gap_tro_se, gap_tro_ci_low, gap_tro_ci_up)

print(result_tro, digits = 3)
```
## Estimate Mean Differences, Standard Errors & CIs for Triangular v. Oval UFOs

```{r}
# Gaps for Triangular versus Oval UFOs

gap_trov <- triangular_UFOs$Y_bar_tr - oval_UFOS$Y_bar_ov

gap_trov_se <- sqrt((triangular_UFOs$sd_tr^2 / triangular_UFOs$n_tr) + (oval_UFOS$sd_ov^2 / oval_UFOS$n_ov))

gap_trov_ci_low <- gap_trov - 1.96 * gap_trov_se

gap_trov_ci_up <- gap_trov + 1.96 * gap_trov_se

result_trov <- cbind(triangular_UFOs[,-1], oval_UFOS[,-(1:2)], gap_trov, gap_trov_se, gap_trov_ci_low, gap_trov_ci_up)

print(result_trov, digits = 3)
```
## Generate Visualization

```{r}
# Generate a plot 
library(ggplot2)

ggplot(avgs, aes(as.factor(Shape), mean_seconds, fill = Shape)) +
  geom_bar(stat = "identity") + 
  labs(y = "Average encounter length (seconds)", x = "UFO Shape Category", title="Mean length of UFO encounter by Craft Shape") + 
  scale_fill_manual(values = c("darkred", "darkorange2", "darkgreen", "darkcyan", "darkorchid4"))

```

As my independent variable is nominal, I figured a bar chart would be the best way to show how the mean encounter length varies for each of the 5 shape categories. This visualization suggests that people have spent the most time encountering round-shaped UFOs, but later data visualizations will show that the round UFO category has a notable outlier. After round UFOs, triangular and oval-shaped crafts have the highest average encounter length. I'll focus on those 3 shape categories rather than all 5. 

## Regression with Dummy Variables

```{r}
# Carry out simple regression analysis using lm(). Generate plot that regresses X variable on Y. Fit a line through observations. 

# Create dummy variables for categorical Shape (x) variable
library(fastDummies)
ufo_sighting_data <- dummy_cols(ufo_sighting_data, select_columns = 'Shape_Category')

ufo_lm <- lm(length_of_encounter_seconds ~ Shape_Category, ufo_sighting_data)

print(ufo_lm)
```

## Plot for Triangular UFOs

```{r}
# Plot distribution for Triangular crafts
Tri_mean <- mean(ufo_sighting_data$length_of_encounter_seconds[ufo_sighting_data$Shape_Category_Triangular == 1])
Ntri_mean <- mean(ufo_sighting_data$length_of_encounter_seconds[ufo_sighting_data$Shape_Category_Triangular == 0])

plot(length_of_encounter_seconds ~ Shape_Category_Triangular, ufo_sighting_data, pch=19, col="steelblue", 
     main="Triangular-Shaped UFOs", 
     ylab="Length of Encounter (Seconds)",
     xlab="Triangular-Shaped Craft versus Non-Triangular Crafts")
points(y = Tri_mean, x=1, col="red", pch=19) 
points(y = Ntri_mean, x=0, col="red", pch=19)
```

## Plot for Oval UFOs

```{r}
# Plot distribution for Oval crafts
Oval_mean <- mean(ufo_sighting_data$length_of_encounter_seconds[ufo_sighting_data$Shape_Category_Oval == 1])
Nov_mean <- mean(ufo_sighting_data$length_of_encounter_seconds[ufo_sighting_data$Shape_Category_Oval == 0])

plot(length_of_encounter_seconds ~ Shape_Category_Oval, ufo_sighting_data, pch=19, col="steelblue", 
     main="Oval-Shaped UFOs", 
     xlab="Oval Crafts versus Non-Oval Crafts",
     ylab="Length of Encounter (Seconds)")

points(y = Oval_mean, x=1, col="red", pch=19)
points(y = Nov_mean, x=0, col="red", pch=19)
```

```{r}
# Plot distribution for Round crafts
Ro_mean <- mean(ufo_sighting_data$length_of_encounter_seconds[ufo_sighting_data$Shape_Category_Round == 1])
Nro_mean <- mean(ufo_sighting_data$length_of_encounter_seconds[ufo_sighting_data$Shape_Category_Round == 0])

plot(length_of_encounter_seconds ~ Shape_Category_Round, ufo_sighting_data, pch=19, col="steelblue",
     main="Round-Shaped UFOs",
     ylab="Length of Encounter (Seconds)",
     xlab="Round vs. Non-Round Crafts")

points(y = Ro_mean, x=1, col="red", pch=19)
points(y = Nro_mean, x=0, col="red", pch=19)
```
This plot allows us to see the outlier point in the round shape category that may influence the relatively high value of its mean when compared to the other shape categories. 